Psych 254 W15 PS #1

### 2015-01-25 analyses/ps1.Rmd
Completed assignment1

**There are several ways in which the data need to be cleaned. One tip-off about this is that a number of variables that should be numeric end up as factors... In particular, we note:**
1. **Some participants appear multiple times (90 rows, but only 87 worker ids)**
2. **There are commas in some of the input.price numbers**
3. **someone entered their response for "dog cost" as a character string: "five hundred"**

@@

**We can conclude that** 
1. **there are three clumps of data, corresponding to the three different objects (sushi, dog, plasma)**, 
2. **recalling that the anchors were set at 9, 2500, and 5000 (remember, we're not yet separating out classes of objects), we see that there seems to be a general trend of under-estimation**

@@

This is a great time to get comfortable with the `%>%` operator. In brief, `%>%` allows you to pipe data from one function to another. So if you would have written:

```{r}
# d <- function(d, other_stuff)
```

you can now write:

```{r}
# d <- d %>% function(other_stufF)
```

That doesn't seem like much, but it's cool when you can replace:

```{r}
# d <- function1(d, other_stuff)
# d <- function2(d, lots_of_other_stuff, more_stuff)
# d <- function3(d, yet_more_stuff)
```

with

```{r}
# d <- d %>% 
#   function1(other_stuff) %>%
#   function2(lots_of_other_stuff, more_stuff) %>%
#   function3(yet_more_stuff)
```

In other words, you get to make a clean list of the things you want to do and chain them together without a lot of intermediate assignments.

@@

```{r}
d.tidy %>% 
  group_by(name, condition) %>% 
  summarise(mean(cost, na.rm = TRUE))
```

@@

```{r}
pcts <- d.tidy %>%
  mutate(pct_change = anchor / cost) %>%
  group_by(name, condition) %>% 
  summarise(mean(cost, na.rm = TRUE))  
```

@@

```{r}
z.scores <- d.tidy %>% ...
```

>>>>

**There are several ways in which the data need to be cleaned. One tip-off about this is that a number of variables that should be numeric end up as factors... In particular, we note:**

1. **Some participants appear multiple times (90 rows, but only 87 worker ids)**

2. **There are commas in some of the input.price numbers**

3. **someone entered their response for "dog cost" as a character string: "five hundred"**

@@

**We can conclude that**

1. **there are three clumps of data, corresponding to the three different objects (sushi, dog, plasma)**,

2. **recalling that the anchors were set at 9, 2500, and 5000 (remember, we're not yet separating out classes of objects), we see that there seems to be a general trend of under-estimation. **

**However, because these three classes of objects are at different scales, it's hard to compare them all on the same axis... In the following plot, we separate out 'dog', 'plasma', and 'sushi', while still lumping together the 'over', 'under', and 'rounded' anchor conditions. This supports our observation that people tend to be under-estimating prices, overall.**

```{r}
g <- (ggplot(d.tidy, aes(x = cost))
      + geom_histogram()
      + facet_wrap(~ name, scales = 'free'))
g
```

@@

This is a great time to get comfortable with the `%>%` operator. In brief, `%>%` allows you to pipe data from one function to another. In other words, you get to make a clean list of the things you want to do and chain them together without a lot of intermediate assignments.

@@

```{r}
d.tidy %>% 
  group_by(name, condition) %>% 
  summarise(mean_cost = mean(cost, na.rm = TRUE))
```

@@

```{r}
pcts <- d.tidy %>%
  mutate(pct_change = abs((anchor - cost) / anchor)) %>%
  group_by(item, condition) %>% 
  summarise(pct = mean(pct_change, na.rm = TRUE))  
pcts
```

@@

```{r}
z.scores <- d.tidy %>% 
  group_by(item) %>%
  mutate(z_item = scale(cost)[,1]) %>%
  group_by(condition, add = TRUE) %>%
  summarise(z = mean(z_item, na.rm = TRUE))
z.scores
```
